Solving Collective Spin Hamiltonians with Decoherence

### 2018-10-01 soc_squeezing/cumulant_expansions.tex
removed redundant notation
fixed a few errors and futher simplified the product of two k-body operators

We now expand a product of basis operators $\S_{\v m}$ and $\S_{\v n}$
into terms that have exactly $s$ spins addressed by both operators:
\begin{align}
  \S_{\v m} \S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v j,\v k;\v m,\v n,s}} P_{\v j} P_{\v k},
  &&
  \sum_{\p{\v j,\v k;\v m,\v n,s}} X \equiv
  \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
      \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X,
  \label{eq:SS_PP}
\end{align}
where the sum over $s$ terminates after $s = \sum_\mu\p{m_\mu+n_\mu}$.
Collecting terms in which $r_{\mu\nu}$ of the $\sigma_\mu$ operators
in $P_{\v j}$ address the same spin as a $\sigma_\nu$ operator in
$P_{\v k}$, we have
\begin{align}
  \S_{\v m}\S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v r,\v J,\v K;\v m,\v n,s}}
  f_{\v m\v n\v r} Q_{\v J} \tilde Q_{\v K},
  \label{eq:SS_PQ}
\end{align}
where $\v J$ and $\v K$ are respectively ordered lists of indices for
spins addressed by one and two single-spin operators in each term of
the expansion for $\S_{\v m} \S_{\v n}$ in \eqref{eq:SS_PP}; the sum
over $\p{\v r,\v J,\v K;\v m,\v n,s}$ implies a sum over all possible
values of $\v r,\v J,\v K$ with the restrictions
\begin{align}
  r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
  &&
  \sum_\nu r_{\mu\nu} \le m_\mu,
  &&
  \sum_\mu r_{\mu\nu} \le n_\nu,
  &&
  \sum_{\mu,\nu} r_{\mu\nu} = s,
\end{align}
\begin{align}
  \abs{\set{J_a^\mu}} = \sum_\mu\p{m_\mu+n_\mu} - 2 s,
  &&
  \abs{\set{K_a^{\mu\nu}}} = s,
  &&
  \set{J_a^\mu} \cap \set{K_a^{\mu\nu}} = \set{};
\end{align}
the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
operators in $P_{\v j}$ and $P_{\v k}$ as prescribed by $\v r$, or
\begin{align}
  f_{\v m\v n\v r}
  &\equiv \sp{\prod_{\mu,\nu}
    { m_\mu - \sum_{\rho>\nu} r_{\mu\rho} \choose r_{\mu\nu} }
    { n_\mu - \sum_{\rho>\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
  \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\rho r_{\mu\rho}}
    \p{n_\mu}_{\sum_\rho r_{\rho\mu}}}
  \sp{\prod_{\mu\nu} r_{\mu\nu}!}^{-1},
\end{align}
where $\p{n}_k\equiv\prod_{j=0}^{k-1}\p{n-j}$ is a falling factorial;
and finally
\begin{align}
  Q_{\v J} \equiv \prod_{\mu,a} \sigma_\mu^{(J_a^\mu)},
  &&
  \tilde Q_{\v K} \equiv \prod_{\mu,\nu,a}
  \sigma_\mu^{(K_a^{\mu\nu})} \sigma_\nu^{(K_a^{\mu\nu})}.
\end{align}

>>>>

We now expand a product of basis operators $\S_{\v m}$ and $\S_{\v n}$
into terms that have exactly $s$ spins addressed by both operators:
\begin{align}
  \S_{\v m} \S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v j,\v k;\v m,\v n,s}} P_{\v j} P_{\v k},
  &&
  \sum_{\p{\v j,\v k;\v m,\v n,s}} X \equiv
  \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
      \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X,
  \label{eq:SS_PP}
\end{align}
where the sum over $s$ terminates after $s = \sum_\mu\p{m_\mu+n_\mu}$.
Collecting terms in which $r_{\mu\nu}$ of the $\sigma_\mu$ operators
in $P_{\v j}$ address the same spin as a $\sigma_\nu$ operator in
$P_{\v k}$, we have
\begin{align}
  \S_{\v m}\S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v r,\v J,\v K;\v m,\v n,s}}
  f_{\v m\v n\v r} P_{\v J} Q_{\v K},
  \label{eq:SS_PQ}
\end{align}
where $\v J$ and $\v K$ are respectively ordered lists of indices for
spins addressed by one and two single-spin operators in each term of
the expansion for $\S_{\v m} \S_{\v n}$ in \eqref{eq:SS_PP}; the sum
over $\p{\v r,\v J,\v K;\v m,\v n,s}$ denotes a sum over all values of
$\v r,\v J,\v K$ with the restrictions
\begin{align}
  r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
  &&
  \sum_\nu r_{\mu\nu} \le m_\mu,
  &&
  \sum_\mu r_{\mu\nu} \le n_\nu,
  &&
  \sum_{\mu,\nu} r_{\mu\nu} = s,
\end{align}
\begin{align}
  \abs{\set{J_a^\mu}}
  = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
  &&
  \abs{\set{K_a^{\mu\nu}}} = r_{\mu\nu},
  &&
  \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \set{};
\end{align}
the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
operators in $P_{\v j}$ and $P_{\v k}$ as prescribed by $\v r$, or
\begin{align}
  f_{\v m\v n\v r}
  &\equiv \sp{\prod_{\mu,\nu}
    { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
    { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
  \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\rho r_{\mu\rho}}
    \p{n_\mu}_{\sum_\rho r_{\rho\mu}}}
  \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1},
\end{align}
where $\p{n}_k\equiv\prod_{j=0}^{k-1}\p{n-j}$ is a falling factorial;
and finally, similarly to the definition of $P_{\v j}$ in
\eqref{eq:SS_PP} we have
\begin{align}
  Q_{\v K} \equiv \prod_{\mu,\nu,a}
  \sigma_\mu^{(K_a^{\mu\nu})} \sigma_\nu^{(K_a^{\mu\nu})}.
\end{align}
Expanding the product
$\sigma_\mu\sigma_\nu\equiv\sum_\rho\eta_{\mu\nu\rho}\sigma_\rho$,
where unlike the sums over $\mu,\nu$ in most of the work above the sum
over $\rho$ here includes an index for the identity operator
$\sigma_0\equiv\1$, we can simplify
\begin{align}
  Q_{\v K}
  = \prod_{\mu,\nu,a} \sum_\rho
  \eta_{\mu\nu\rho} \sigma_\rho^{(K_a^{\mu\nu})},
\end{align}
and collect terms for each $\p{\mu,\nu}$ that are unique up to a
permutation of indices and have $c_\rho$ single-spin operators of
the type $\sigma_\rho$, giving us
\begin{align}
  Q_{\v K}
  \cong \prod_{\mu,\nu} \sum_{\sum_\lambda c_\lambda=r_{\mu\nu}}
  g^{\mu\nu}_{\v c\v r} \prod_\rho \p{\eta_{\mu\nu\rho}}^{c_\rho}
  \prod_{a=1}^{c_\rho} \sigma_\rho^{(K^{\mu\nu\rho}_{\v ca})},
\end{align}
where $\cong$ denotes equality under the sum in \eqref{eq:SS_PQ} (as
we have freely permuted spin indices and introduced explicit
dependence on $\v r$ for convenience); the factor
$g^{\mu\nu}_{\v c\v r}$ counts the number of ways to choose operators
in $Q_{\v K}$ according to $\v c$ (i.e.~the number of choices that are
equivalent up to permutation of indices in $\v K$), or
\begin{align}
  g^{\mu\nu}_{\v c\v r} \equiv \prod_\rho
  { r_{\mu\nu} - \sum_{\lambda<\rho} c_\lambda \choose c_\rho }
  = \f{r_{\mu\nu}!}{\prod_\rho c_\rho!};
\end{align}
and
\begin{align}
  K^{\mu\nu\rho}_{\v ca}
  \equiv K^{\mu\nu}_{\sum_{\lambda<\rho}c_\lambda+a}
\end{align}
indexes the $a$-th operator of type $\sigma_\rho$, given a choice of
$\v c$.
